### 2012-01-30 trunk/GUIErrorDetector/paper/implementation.tex
update

We implemented our technique using the WALA framework~\cite{walatutorial}.
Our prototype tool takes Java bytecode as inputs, works in a fully
automatic and push-button way to detect potential invalid thread access
errors, and scales to realistic programs.  
Optionally, the user may specify additional heuristics to guide call graph
traverse, and annotate calling relations for native methods (Section~\ref{sec:heuristic}).

Our current implementation supports four widely-used GUI frameworks, namely
SWT, Eclipse plugin development environment, Java Swing, and Android.
Our implementation handles all of the Java language, and also takes
all dependent libraries (e.g., \CodeIn{rt.jar} and \CodeIn{android.jar})
into analysis. Since Android applications are often fully encrypted and shipped in Dalvik
bytecode as a single apk file, our implementation first uses
android-apktool\footnote{\url{http://code.google.com/p/android-apktool/}} to
decrypt the apk file, and then uses the 
ded translator~\cite{Enck:2011:SAA:2028067.2028088} to convert
Dalvik bytecode to Java bytecode before feeding to WALA.  The Android system
library (i.e., \CodeIn{android.jar}) uses many ``stub'' classes as
placeholder for the sake of efficiency. We manually re-compiled the whole
\CodeIn{android.jar} from its source code by replacing all ``stub'' classes
with real class files.

>>>>

We implemented the proposed technique using the WALA framework~\cite{walatutorial}.
Our prototype tool takes Java bytecode as inputs, works in a fully
automatic and push-button way to detect potential invalid thread access
errors, and scales to realistic programs.  
Optionally, users may specify additional heuristic filters (Section~\ref{sec:heuristic}) %to guide call graph traverse
, and annotate call relationships for native methods (Section~\ref{sec:annotation}).

Our current implementation supports four widely-used GUI frameworks, namely
SWT, Eclipse plugin development, Java Swing, and Android.
Our implementation %handles all of the Java language, and
also takes
all dependent libraries (e.g., \CodeIn{rt.jar} and \CodeIn{android.jar})
into analysis. We implemented the reflection-aware call graph construction
 algorithm (Section~\ref{sec:cg})
using WALA's \textit{bypass logic}. Our tool parses the configuration file
in an Android application and modifies the bytecode to replace all
reflection calls with object creation instructions.
Since Android applications are often fully encrypted and shipped in Dalvik
bytecode as a single apk file, our tool first uses
android-apktool~\cite{apktool} to
decrypt the apk file, and then uses the 
ded translator~\cite{Enck:2011:SAA:2028067.2028088} to convert
Dalvik bytecode to Java bytecode before feeding to WALA.  The Android system
library (i.e., \CodeIn{android.jar}) uses many ``stub'' classes as
placeholder for the sake of efficiency. We manually re-compiled the whole
\CodeIn{android.jar} from its source code by replacing all ``stub'' classes
with real class files.

%Static Single Assignment (SSA) transformation to get limited flow sensitivity

### 2012-01-30 trunk/GUIErrorDetector/paper/introduction.tex
update

\item The single-thread rule incurs less overhead.
A GUI framework that attempts to carefully lock critical sections can spend a substantial
amount of time and space managing locks. Whenever the framework calls a method that might
be implemented in client code (for example, any public or protected method in a public class),
the framework must save its state and release all locks so that the client code can grab locks
if necessary. When GUI objects return from the method, the framework must re-grab their locks and
restore states. Therefore, all applications using the framework will bear the cost of this, even though most
applications do not require concurrent access to the GUI.

@@

\end{alltt}
\end{CodeOut}
\vspace*{-2.0ex} \Caption{{\label{fig:androiderror} A real
bug reported on 11/17/2011 in Mozilla firefox release 10.0.a.rc3
for the Android platform (bugzilla ID: 703256). \CodeIn{LayerView}'s constructor
calls method \CodeIn{setRender} that
spawns a new thread at line 283. The new thread calls back method
\CodeIn{onSurfaceChanged} that access GUI objects at line 222, and
causes an invalid thread access error.
Our tool
can find this error and generate a warning as shown in
Figure~\ref{fig:report}.
}} %\vspace{-5mm}
\end{figure}

@@

Invalid thread access errors have significantly affect the
usability of GUIs, but they have not been thoroughly studied in
the context of multithreaded GUI applications. To ensure GUIs
behave correctly, effective techniques must be developed to alleviate this problem.

@@

A standard way to fix an
invalid thread access error is to replace direct GUI object accessing code 
with asynchronized message passing, to ensure a GUI object is accessed in
the UI thread. For example, to fix the bug in Figure~\ref{fig:androiderror},
a developer would wrap line 222 inside a \CodeIn{post} message passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is a
a standard way to send messages to the UI-thread.}, as follows:

@@

One may wonder that a simple analysis to check whether every GUI object
accessing code has been wrapped by (a)synchronized message passing may
be capable to find invalid thread errors. However, such an approach is problematic,
because not every GUI object access needs to be wrapped by a message passing
method, e.g., methods that are only called from UI-thread can safely
access any GUI object. On the other hand, using asynchronized message passing has no timing
guarantee. It is entirely possible that, in some cases, a GUI object has already been
disposed before the message sent to arrives, causing other bugs. 
Thus, programmers can not use it everywhere to wrap all GUI object accessing code.
In our evaluation on real world programs, we found this simple analysis 
can not be used in practice: it issued
an unacceptable number of false positives and redundant warnings.

@@

Our static analysis is independent of the call graph construction algorithms.
Given a sound call graph, our analysis is sound in that it does
not miss true positives. We note, however, that computing a sound
call graph with acceptable precision for GUI applications in the presence of reflection
is non-trivial.
Existing call graph construction algorithms such as RTA~\cite{rta}
and k-CFA~\cite{kcfa}
fails to build a sufficiently complete call graph, due to their limitations
to reason about reflections.  To alleviate this problem,
we present a method to build reflection-aware call graph, and
also compare its usefulness with existing call graph construction algorithms in
our experiments. We also note that static analysis may report false positives due to its
conservative nature, or multiple warnings that actually correspond
to the same error. To alleviate such limitations, we devise a set
 of heuristics to filter likely false positives and redundant warnings.

@@

We instantiated our technique for four popular GUI frameworks namely SWT,
eclipse plugin environment, Java Swing, and Android. 
Swing and SWT are unquestionably 
two dominating GUI frameworks in developing desktop Java applications.
Eclipse and its plugin development environment is the de-facto standard
IDE for Java. Android is the No. 1 platform for
developing mobile applications with market share 56\% as of Sep 2011.
Although our technique is generally applicable for GUI
frameworks adapting the single-thread rule, a specific framework
may have its own treatment of \textit{UI thread} and \textit{call graph
entry nodes}. Thus, we must explicitly identify those framework-specific
parts in our prototype tool implementation.

@@

We implement an invalid thread error detection tool that supports
the above four GUI frameworks. Our tool works in a fully-automatic,
push-button manner, and scales to realistic programs.
We evaluate our implementation on 10 subjects from four different
frameworks over 75000 LOC. The experiment results demonstrate
that: our technique is effective (it found 12 real-world errors
on all supported frameworks, and produced significantly less
false positives and redundant warnings than the simple analysis);
our reflection-aware call graph construction algorithm is
critical in finding errors in Android applications,
and our proposed
heuristics are useful (it significantly reduces the number of
false positives and redundant warnings).

@@

\item \textbf{Technique.} We formulate finding
the invalid thread access error as a call graph reachability problem,
and present a general error detection technique.
In particular, to deal with the unique features in
multithreaded GUI programs, we propose a reflection-aware
call graph construction method. (Section~\ref{sec:technique}).

@@

\item \textbf{Evaluation.} We performed an experiment on 10 subjects
from 4 different platforms over 75000 LOC to show the usefulness
of the proposed technique (Section~\ref{sec:evaluation}).

>>>>

\item The single-thread rule incurs less overhead.
A GUI framework that attempts to carefully lock critical sections can spend a substantial
amount of time and space managing locks. Whenever the framework calls a method that might
be implemented in client code (e.g., any public or protected method in a public class),
the framework must save its state and release all locks so that the client code can grab locks
if necessary. When GUI objects return from the method, the framework must re-grab their locks and
restore states. Therefore, all applications using the framework will bear the cost of this, even though most
applications do not require concurrent access to the GUI.

@@

\end{alltt}
\end{CodeOut}
\vspace*{-2.0ex} \Caption{{\label{fig:androiderror} A real
bug reported on 11/17/2011 in Mozilla firefox release 10.0.a.rc3
for the Android platform (bugzilla ID: 703256). \CodeIn{LayerView}'s constructor
calls method \CodeIn{setRender} that
spawns a new thread at line 283. The new thread calls back method
\CodeIn{onSurfaceChanged} that access GUI objects at line 222, and
causes an invalid thread access error.
Our tool
can find this error and generate a report as shown in
Figure~\ref{fig:report}.
}} %\vspace{-5mm}
\end{figure}

@@

Invalid thread access errors have significantly affected the
usability of GUIs, but they have not been thoroughly studied in
the context of multithreaded GUI applications. To ensure GUIs
behave correctly, effective techniques must be developed to alleviate this problem.

@@

A standard way to fix an
invalid thread access error is to replace direct GUI object accessing code 
with (a)synchronized message passing, to ensure a GUI object is accessed in
the UI thread. For example, to fix the bug in Figure~\ref{fig:androiderror},
a developer would wrap line 222 inside a \CodeIn{post} message passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is a
a standard way to send asynchronized messages to the UI-thread.}, as follows:

@@

One may wonder that a simple analysis to check whether every GUI object
accessing code has been wrapped by (a)synchronized message passing may
be capable to find invalid thread errors. However, such an approach is problematic,
because not every GUI object access needs to be wrapped by message passing
method. For example, methods that are only called from UI-thread can safely
access any GUI object. On the other hand, using asynchronized message passing has no timing
guarantee. It is entirely possible that, in some cases, a GUI object has already been
disposed before the message sent to arrives, causing other bugs. 
Thus, programmers can not use it everywhere to wrap all GUI object accessing code.
In our evaluation on real world programs, we found this simple analysis 
can not be used in practice: it issued
an unacceptable number of invalid warnings. %false positives and redundant warnings.

@@

Our static analysis is independent of the call graph construction algorithms.
Given a sound call graph, our analysis is sound in that it does
not miss true positives. We note, however, that computing a sound
call graph with acceptable precision for GUI applications in the presence of reflection
is non-trivial.
Existing call graph construction algorithms such as RTA~\cite{rta}
and k-CFA~\cite{kcfa}
fails to build a good call graph, due to their limitations
to reason about reflections.  To alleviate this problem,
we present an algorithm to build reflection-aware call graph, and
also compare its usefulness with existing call graph construction algorithms in
our experiments. We also note that static analysis may report false positives due to its
conservative nature, or multiple warnings that actually correspond
to the same error. To address such limitations, we devise a set
 of heuristics to filter likely false positives and redundant warnings.

@@

We instantiated our technique for four popular GUI frameworks namely SWT,
Eclipse plugin environment, Java Swing, and Android. 
Swing and SWT are unquestionably 
two dominating GUI frameworks in developing desktop Java applications.
Eclipse and its plugin development environment is the de-facto standard
IDE for Java. Android is the No. 1 platform for
developing mobile applications with market share 56\% as of Sep 2011.
Although our technique is generally applicable for GUI
frameworks adapting the single-thread rule, a specific framework
may have its own treatment of \textit{UI thread} and \textit{program 
entry points}. Thus, we must explicitly identify those framework-specific
parts in our prototype tool implementation.

@@

We implement an invalid thread error detection tool that supports
the above four GUI frameworks. Our tool works in a fully-automatic
 manner, and scales to realistic programs.
We evaluate our implementation on 9 subjects from four different
frameworks over 67KLOC with 1.4MLOC of dependent library code.
The experiment results demonstrate
that: our technique is effective (it found 11 real-world errors
from four supported frameworks
%, and produced significantly less
%false positives and redundant warnings than the simple analysis
);
our reflection-aware call graph construction algorithm is
critical in finding errors in Android applications,
and our proposed
heuristics are useful (it significantly reduces the number of
false positives and redundant warnings).

@@

\item \textbf{Technique.} We formulate finding
the invalid thread access error as a call graph reachability problem,
and present a general error detection technique.
In particular, to deal with the unique features in
multithreaded GUI programs, we propose a reflection-aware
call graph construction algorithm. (Section~\ref{sec:technique}).

@@

\item \textbf{Evaluation.} We performed an experiment on 9 subjects
from 4 different platforms over 67000 LOC to show the usefulness
of the proposed technique (Section~\ref{sec:evaluation}).
